Weighted Model Counting with Conditional Measures

### 2020-07-26 paper.tex
Some writing and editing

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{complexity}
\usepackage{soul}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{subcaption}
\usepackage{multirow}

@@

\begin{itemize}
\item Describe \cref{tbl:notation_example} as an example.
\item[F] Feedback: explain with examples: models $=$ atoms of algebra, formulas
  $=$ all elements.
\end{itemize}

@@

\begin{proposition}
  Every weight function induces a measure:
  \[
    M_W(x) = \begin{cases}
      0 & \text{if } x = \bot \\
      W(u) & \text{if } x = \{ u \} \\
      \sum_{\{u\} \le x} W(u) & \text{otherwise.}
    \end{cases}
  \]
\end{proposition}
The process of calculating the value of $M_W(x)$ for some $x \in 2^{2^U}$ with a
given definition of $W$ is known as \emph{weighted model counting}.
\begin{proof}
  \todo[inline]{Prove that $M_W$ satisfies the definition of a measure.}
\end{proof}

@@

\begin{example}[Relation to the classical view of WMC] \label{example:construction}
  Let $\mathcal{L}$ be a propositional logic with a set of atoms $U = \{ p, q
  \}$. Let $w : \{ p, q, \neg p, \neg q \} \to \mathbb{R}_{\ge 0}$ be the
  \emph{weight function} defined by
  \begin{align*}
    w(p) = 0.3, \quad w(\neg p) = 0.7, \quad w(q) = 0.2, \quad w(\neg q) = 0.8.
  \end{align*}
  Let $\Delta$ be a theory in $\mathcal{L}$ with a sole axiom $p$. Then
  $\Delta$ has two models, i.e., $\{ p, q \}$ and $\{ p, \neg q \}$. The
  \emph{weighted model count} (WMC) \cite{DBLP:journals/ai/ChaviraD08} of
  $\Delta$ is then
  \begin{equation} \label{eq:wmc_example}
    \sum_{\omega \models \Delta} \prod_{\omega \models l} w(l) =
    w(p)w(q) + w(p)w(\neg q) = 0.3.
  \end{equation}

  Alternatively, we can define $W_p\colon 2^{\{p\}} \to \mathbb{R}_{\ge 0}$ as
  \[
    W_p(\{ p \}) = 0.3, \quad W_p(\emptyset) = 0.7
  \]
  and $W_q\colon 2^{\{q\}} \to \mathbb{R}_{\ge 0}$ as
  \[
    W_q(\{ q \}) = 0.2, \quad W_q(\emptyset) = 0.8.
  \]
  Let $M$ be the measure on $2^{2^U}$ induced by $W = W_p \cdot W_q$. Then,
  equivalently to \cref{eq:wmc_example}, we can write
  \[
    M(p) = W(\{ p, q \}) + W(\{ p, \neg q \}) = W_p(\{p\})W_q(\{q\}) +
    W_p(\{p\})W_q(\{\neg q\}) = 0.3.
  \]
\end{example}

@@

\todo[inline]{Include this in the example.}
Given a theory $\Delta$ in a logic $\mathcal{L}$, the usual way of using WMC to
compute the probability of a query $q$ is
\cite{DBLP:conf/uai/Belle17,DBLP:conf/aaai/SangBK05}
\[
  \Pr_{\Delta, w}(q) = \frac{\WMC_w(\Delta \land q)}{\WMC_w(\Delta)}.
\]
In our algebraic formulation, this can be computed...

@@

\todo[inline]{This theorem needs a special case for zero weights.}
\begin{theorem}
  A measure $M\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$ is factorable if and only
  if
  \begin{equation} \label{eq:wmccondition}
  M(u \land v) = M(u)M(v)
  \end{equation}
  for all distinct $u, v \in U \cup \{ \neg w \mid w \in U \}$ such that $u \ne
  \neg v$.
\end{theorem}
\begin{proof}
  ($\Leftarrow$) For each $x \in U$, let $W_x\colon 2^{\{x\}} \to
  \mathbb{R}_{\ge 0}$ be defined by $W_x(\{ x \}) = M(x)$ and $W_x(\emptyset) =
  M(\neg x)$. Let $M_W$ be the measure induced by
  \[
    W = \prod_{x \in U} W_x.
  \]
  We will show that $M = M_W$. First, note that $M_w(\bot) = 0 = M(\bot)$ by the
  definitions of both $M_w$ and $M$. Second, let
  \begin{equation} \label{eq:def_of_a}
    a = \bigwedge_{u \in U} a_u
  \end{equation}
  be an atom in $2^{2^U}$ such that $a_u \in \{ u, \neg u \}$ for all $u \in U$.
  Then
  \[
    M_W(a) = \prod_{u \in U} W_u(a_u) = \prod_{u \in U} M(a_u) = M
    \left(\bigwedge_{u \in U} a_u \right) = M(a)
  \]
  \todo{by what?}
  by \cref{def:wmc,eq:wmccondition,eq:def_of_a}. Finally, note that if $M_W$ and
  $M$ agree on all atoms, then they must also agree on all other non-zero
  elements of the Boolean algebra.

@@

\begin{itemize}
\item Encodings:
  \begin{itemize}
  \item \texttt{d02} \cite{DBLP:conf/kr/Darwiche02}
  \item \texttt{sbk05} \cite{DBLP:conf/aaai/SangBK05}
  \item \texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05}
  \item \texttt{cd06} \cite{DBLP:conf/sat/ChaviraD06} (supposed to be the best)
  \item \texttt{db20} (mine)
  \end{itemize}
\item \texttt{cd05} relaxes the encoding so much that extra models become
  possible. They are supposed to be filtered out by the algorithm, but mine
  can't do that because it doesn't deal with models. Same for \texttt{cd06}
  because it's based on \texttt{cd05}.
\item \texttt{sbk05} uses my trick with dividing probabilities. That could
  explain small inaccuracies in its answers.
\end{itemize}

\subsection{Algebraic Decision Diagrams and ADDMC}

References
\begin{itemize}
\item ADDs \cite{DBLP:journals/fmsd/BaharFGHMPS97}
\item background reading
  \begin{itemize}
  \item Compiling Bayesian Networks Using Variable Elimination (Chavira and
    Darwiche) \cite{DBLP:conf/ijcai/ChaviraD07}
  \item On the Relationship between Sum-Product Networks and Bayesian Networks
    (Zhao et al.) \cite{DBLP:conf/icml/ZhaoMP15}
  \end{itemize}
\end{itemize}

@@

\begin{itemize}

\item We assume that all variables in the Bayesian network have at least two
  values.
\item Have an example of how the ADDs function in this situation. If not for the
  paper, then at least for slides. Use the framework to check its correctness.
\item The function $\phi$ created by the algorithm can be seen as a measure on
  the BA $2^{2^U}$.
\item We extend the Gaifman graph to add edges when two variables occur in the
  same CPT (e.g., including the edge from $A$ to $B$ when the CPT is $\Pr(A \mid
  B)$).
\end{itemize}

@@

\begin{algorithm}
  $\phi \gets 1$\;
  \For{$X \in V$}{
    \textit{let} $\mathrm{pa}(X) = \{ Y_1, \dots, Y_n \}$\;
    $\mathrm{CPT}_X \gets 0$\;
    \eIf{$|\im X| = 2$}{
      \For{$(y_1, \dots, y_n) \in \prod_{i = 1}^n \im Y_i$}{
        $p_1 \gets \Pr(X = 1 \mid Y_1 = y_1, \dots, Y_n = y_n)$\;
        $p_0 \gets \Pr(X \ne 1 \mid Y_1 = y_1, \dots, Y_n = y_n)$\;
        $\mathrm{CPT}_X \gets \mathrm{CPT}_X + p_1[\lambda_{X=1}] \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}] + p_0 \overline{[\lambda_{X=1}]} \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}]$\;
      }
    }{
      \textit{let} $\im X = \{ x_1, \dots, x_m \}$\;
      \For{$x \in \im X$ {\rm \textbf{and}} $(y_1, \dots, y_n) \in \prod_{i =
          1}^n \im Y_i$}{
        $p_x \gets \Pr(X = x \mid Y_1 = y_1, \dots, Y_n = y_n)$\;
        $\mathrm{CPT}_X \gets \mathrm{CPT}_X + p_x[\lambda_{X=x}] \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}] + \overline{[\lambda_{X=1}]} \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}]$\;
      }
      $\mathrm{CPT}_X \gets \mathrm{CPT}_X \cdot \left( \sum_{i=1}^m [\lambda_{X
          = x_i}] \right) \cdot \prod_{i=1}^m \prod_{j=i+1}^m
      (\overline{[\lambda_{X = x_i}]} + \overline{[\lambda_{X = x_j}]})$\;
    }
    $\phi \gets \phi \cdot \mathrm{CPT}_X$\;
  }
  \Return{$\phi$}\;
\end{algorithm}

@@

\begin{example}
  \begin{figure}
    \centering
    \begin{subfigure}{0.2\textwidth}
      \centering
      \begin{tikzpicture}
        \node[latent] (W) {W};
        \node[latent, below right=of W] (T) {T};
        \node[latent, below left=of W] (F) {F};
        \edge {W} {F};
        \edge {W} {T};
      \end{tikzpicture}
    \end{subfigure}%
    \begin{subfigure}{0.8\textwidth}
      \centering
      \begin{tabular}[t]{cc}
        \toprule
        $w$ & $\Pr(W = w)$ \\
        \midrule
        1 & 0.5 \\
        0 & 0.5 \\
        \bottomrule
      \end{tabular}
      \begin{tabular}[t]{ccc}
        \toprule
        $w$ & $f$ & $\Pr(F = f \mid W = w)$ \\
        \midrule
        1 & 1 & 0.6 \\
        1 & 0 & 0.4 \\
        0 & 1 & 0.1 \\
        0 & 0 & 0.9 \\
        \bottomrule
      \end{tabular}
      \begin{tabular}[t]{ccc}
        \toprule
        $w$ & $t$ & $\Pr(T = t \mid W = w)$ \\
        \midrule
        1 & $l$ & 0.2 \\
        1 & $m$ & 0.4 \\
        1 & $h$ & 0.4 \\
        0 & $l$ & 0.6 \\
        0 & $m$ & 0.3 \\
        0 & $h$ & 0.1 \\
        \bottomrule
      \end{tabular}
    \end{subfigure}
    \caption{An example Bayesian network with its CPTs}
    \label{fig:example_bn}
  \end{figure}

  The Bayesian network in \cref{fig:example_bn} has
  \begin{align*}
    V &= \{ W, F, T \}, \\
    \mathrm{pa}(W) &= \emptyset, \\
    \mathrm{pa}(F) &= \mathrm{pa}(T) = \{ W \}, \\
    \im W &= \im F = \{ 0, 1 \}, \\
    \im T &= \{ l, m, h \}, \\
    E(W) &= \{ \lambda_{W=1} \}, \\
    E(F) &= \{ \lambda_{F=1} \}, \\
    E(T) &= \{ \lambda_{T=l}, \lambda_{T=m}, \lambda_{T=h} \}, \\
    E^*(W) &= \{ \lambda_{W=1} \}, \\
    E^*(F) &= \{ \lambda_{F=1}, \lambda_{W=1} \}, \\
    E^*(T) &= \{ \lambda_{T=l}, \lambda_{T=m}, \lambda_{T=h}, \lambda_{W=1} \}, \\
    \mathrm{CPT_W} &= 0.5[\lambda_{W=1}]+0.5\overline{[\lambda_{W=1}]} = 0.5 \cdot 1, \\
    \mathrm{CPT_F} &= 0.6[\lambda_{F=1}] \cdot [\lambda_{W=1}] + 0.4[\lambda_{F=0}] \cdot [\lambda_{W=1}] + 0.1[\lambda_{F=1}] \cdot [\lambda_{W=0}] + 0.9[\lambda_{F=0}] \cdot [\lambda_{W=0}] \\
      &= 0.6[\lambda_{F=1}] \cdot [\lambda_{W=1}] + 0.4\overline{[\lambda_{F=1}]} \cdot [\lambda_{W=1}] + 0.1[\lambda_{F=1}] \cdot \overline{[\lambda_{W=1}]} + 0.9\overline{[\lambda_{F=1}]} \cdot \overline{[\lambda_{W=1}]}, \\
    \mathrm{CPT_T} &= ([\lambda_{T=l}] + [\lambda_{T=m}] + [\lambda_{T=h}]) \cdot (\overline{[\lambda_{T=l}]} + \overline{[\lambda_{T=m}]}) \cdot (\overline{[\lambda_{T=l}]} + \overline{[\lambda_{T=h}]}) \cdot (\overline{[\lambda_{T=m}]} + \overline{[\lambda_{T=h}]}) \cdot (\dots),
  \end{align*}
  and can be encoded in a DIMACS-like CNF format as
  \[
    \begin{array}{l r r l l}
      \lambda\sb{T=l} &\lambda\sb{T=m} &\lambda\sb{T=h} & &0 \\
                      &-\lambda\sb{T=l} &-\lambda\sb{T=m} & &0 \\
                      &-\lambda\sb{T=l} &-\lambda\sb{T=h} & &0 \\
                      &-\lambda\sb{T=m} &-\lambda\sb{T=h} & &0 \\
      w &\lambda\sb{W=1} & &0.5 &0.5 \\
      w &\lambda\sb{F=1} &\lambda\sb{W=1} &0.6 &0.4 \\
      w &\lambda\sb{F=1} &-\lambda\sb{W=1} &0.1 &0.9 \\
      w &\lambda\sb{T=l} &\lambda\sb{W=1} &0.2 &1 \\
      w &\lambda\sb{T=m} &\lambda\sb{W=1} &0.4 &1 \\
      w &\lambda\sb{T=h} &\lambda\sb{W=1} &0.4 &1 \\
      w &\lambda\sb{T=l} &\lambda\sb{W=0} &0.6 &1 \\
      w &\lambda\sb{T=m} &\lambda\sb{W=0} &0.3 &1 \\
      w &\lambda\sb{T=h} &\lambda\sb{W=0} &0.1 &1
    \end{array}
  \]
  with each $\lambda$ replaced with a unique positive integer.
\end{example}

The last two numbers are the positive and the negative probabilities,
respectively---sometimes they add to one, and sometimes the negative probability
is one, regardless of the value of the first probability.

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{complexity}
\usepackage{soul}
\usepackage{subcaption}
\usepackage{multirow}

@@

\todo[inline]{Describe \cref{tbl:notation_example} as an example.}
\todo[inline]{Feedback: explain with examples: models $=$ atoms of algebra,
  formulas $=$ all elements.}

@@

\begin{proposition} \label{prop:measure}
  Every weight function induces a measure, i.e, for any weight function $W\colon
  2^U \to \mathbb{R}_{\ge 0}$, the function $M_W\colon 2^{2^U} \to
  \mathbb{R}_{\ge 0}$, defined as
  \[
    M_W(x) = \begin{cases}
      0 & \text{if } x = \bot \\
      W(u) & \text{if } x = \{ u \} \\
      \sum_{\{u\} \le x} W(u) & \text{otherwise,}
    \end{cases}
  \]
  is a measure.
\end{proposition}
\begin{proof}
  \todo[inline]{Prove that $M_W$ satisfies the definition of a measure.}
\end{proof}

The process of calculating the value of $M_W(x)$ for some $x \in 2^{2^U}$ with a
given definition of $W$ is known as \emph{weighted model counting}.

@@

\begin{example}[Relation to the classical view of WMC] \label{example:construction}
  Let $\mathcal{L}$ be a propositional logic with a set of atoms $U = \{ p, q
  \}$. Let $w : \{ p, q, \neg p, \neg q \} \to \mathbb{R}_{\ge 0}$ be the
  \emph{weight function} defined by
  \begin{align*}
    w(p) = 0.3, \quad w(\neg p) = 0.7, \quad w(q) = 0.2, \quad w(\neg q) = 0.8.
  \end{align*}
  Let $\Delta$ be a theory in $\mathcal{L}$ with a sole axiom $p$. Then
  $\Delta$ has two models, i.e., $\{ p, q \}$ and $\{ p, \neg q \}$. The
  \emph{weighted model count} (WMC) \cite{DBLP:journals/ai/ChaviraD08} of
  $\Delta$ is then
  \begin{equation} \label{eq:wmc_example}
    \mathrm{WMC}(\Delta) = \sum_{\omega \models \Delta} \prod_{\omega \models l} w(l) = w(p)w(q) + w(p)w(\neg q) = 0.3.
  \end{equation}

  Alternatively, we can define $W_p\colon 2^{\{p\}} \to \mathbb{R}_{\ge 0}$ as
  \[
    W_p(\{ p \}) = 0.3, \quad W_p(\emptyset) = 0.7
  \]
  and $W_q\colon 2^{\{q\}} \to \mathbb{R}_{\ge 0}$ as
  \[
    W_q(\{ q \}) = 0.2, \quad W_q(\emptyset) = 0.8.
  \]
  Let $M$ be the measure on $2^{2^U}$ induced by $W = W_p \cdot W_q$. Then,
  equivalently to \cref{eq:wmc_example}, we can write
  \[
    M(p) = W(\{ p, q \}) + W(\{ p \}) = W_p(\{p\})W_q(\{q\}) +
    W_p(\{p\})W_q(\emptyset) = 0.3.
  \]

  Given the theory $\Delta$, we can compute the probability of a query $q$ as
  \cite{DBLP:conf/uai/Belle17,DBLP:conf/aaai/SangBK05}
  \[
    \Pr_{\Delta, w}(q) = \frac{\WMC(\Delta \land q)}{\WMC(\Delta)}.
  \]
  The same thing can be accomplished using the algebraic formulation, with $M$
  replacing $\mathrm{WMC}$.
\end{example}

@@

@@

\begin{theorem}
  A measure $M\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$ is factorable if and only
  if
  \begin{equation} \label{eq:wmccondition}
  M(u \land v) = M(u)M(v)
  \end{equation}
  for all distinct $u, v \in U \cup \{ \neg w \mid w \in U \}$ such that $u \ne
  \neg v$.
\end{theorem}
\begin{proof}
  \todo[inline]{Check if the proof is valid for zero weights.}
  ($\Leftarrow$) For each $x \in U$, let $W_x\colon 2^{\{x\}} \to
  \mathbb{R}_{\ge 0}$ be defined by $W_x(\{ x \}) = M(x)$ and $W_x(\emptyset) =
  M(\neg x)$. Let $M_W$ be the measure induced by
  \[
    W = \prod_{x \in U} W_x.
  \]
  We will show that $M = M_W$. First, note that $M_w(\bot) = 0 = M(\bot)$ by
  \cref{def:wmc,prop:measure}. Second, let
  \[
    a = \bigwedge_{u \in U} a_u
  \]
  be an atom in $2^{2^U}$ such that $a_u \in \{ u, \neg u \}$ for all $u \in U$.
  Then
  \[
    M_W(a) = W(\{ a_u \mid u \in U \}) = \prod_{u \in U} W_u(\{a_u\}) = \prod_{u
      \in U} M(a_u) = M \left(\bigwedge_{u \in U} a_u \right) = M(a)
  \]

  Finally, note that if $M_W$ and
  $M$ agree on all atoms, then they must also agree on all other non-zero
  elements of the Boolean algebra.

@@

\begin{itemize}
\item Encodings:
  \begin{itemize}
  \item \texttt{d02} \cite{DBLP:conf/kr/Darwiche02}
  \item \texttt{sbk05} \cite{DBLP:conf/aaai/SangBK05}
  \item \texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05}
  \item \texttt{cd06} \cite{DBLP:conf/sat/ChaviraD06} (supposed to be the best)
  \item \texttt{db20} (mine)
  \end{itemize}
\item \texttt{cd05} relaxes the encoding so much that extra models become
  possible. They are supposed to be filtered out by the algorithm, but mine
  can't do that because it doesn't deal with models. Same for \texttt{cd06}
  because it's based on \texttt{cd05}.
\end{itemize}

\subsection{Algebraic Decision Diagrams and Their Use in Probabilistic
  Inference}

Cover:
\begin{itemize}
\item ADDs provide an efficient way to manipulate functions from BAs/power sets
  \cite{DBLP:journals/fmsd/BaharFGHMPS97}.
\item background reading
  \begin{itemize}
  \item Compiling Bayesian Networks Using Variable Elimination (Chavira and
    Darwiche) \cite{DBLP:conf/ijcai/ChaviraD07}
  \item On the Relationship between Sum-Product Networks and Bayesian Networks
    (Zhao et al.) \cite{DBLP:conf/icml/ZhaoMP15}
  \end{itemize}
\end{itemize}

@@

@@

\begin{algorithm}
  \caption{Encoding a Bayesian network as a function $2^U \to \mathbb{R}_{\ge
      0}$}
  \label{alg:encoding}
  \KwData{a Bayesian network with vertices $V$ and probability distribution
    $\Pr$}
  \KwResult{a function $\phi\colon 2^U \to \mathbb{R}_{\ge 0}$}
  $\phi \gets 1$\;
  \For{$X \in V$}{
    \textit{let} $\mathrm{pa}(X) = \{ Y_1, \dots, Y_n \}$\;
    $\mathrm{CPT}_X \gets 0$\;
    \eIf{$|\im X| = 2$}{
      \For{$(y_1, \dots, y_n) \in \prod_{i = 1}^n \im Y_i$}{
        $p_1 \gets \Pr(X = 1 \mid Y_1 = y_1, \dots, Y_n = y_n)$\;
        $p_0 \gets \Pr(X \ne 1 \mid Y_1 = y_1, \dots, Y_n = y_n)$\;
        $\mathrm{CPT}_X \gets \mathrm{CPT}_X + p_1[\lambda_{X=1}] \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}] + p_0 \overline{[\lambda_{X=1}]} \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}]$\;
      }
    }{
      \textit{let} $\im X = \{ x_1, \dots, x_m \}$\;
      \For{$x \in \im X$ {\rm \textbf{and}} $(y_1, \dots, y_n) \in \prod_{i =
          1}^n \im Y_i$}{
        $p_x \gets \Pr(X = x \mid Y_1 = y_1, \dots, Y_n = y_n)$\;
        $\mathrm{CPT}_X \gets \mathrm{CPT}_X + p_x[\lambda_{X=x}] \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}] + \overline{[\lambda_{X=1}]} \cdot
        \prod_{i=1}^n [\lambda_{Y_i=y_i}]$\;
      }
      $\mathrm{CPT}_X \gets \mathrm{CPT}_X \cdot \left( \sum_{i=1}^m [\lambda_{X
          = x_i}] \right) \cdot \prod_{i=1}^m \prod_{j=i+1}^m
      (\overline{[\lambda_{X = x_i}]} + \overline{[\lambda_{X = x_j}]})$\;
    }
    $\phi \gets \phi \cdot \mathrm{CPT}_X$\;
  }
  \Return{$\phi$}\;
\end{algorithm}

\todo[inline]{Describe \cref{alg:encoding}}

\subsection{Proof of Correctness}

@@

\subsection{An Example}

\todo[inline]{Reformulate this as a running example throughout the section.}

\begin{figure}
  \centering
  \begin{subfigure}{0.2\textwidth}
    \centering
    \begin{tikzpicture}[edge from parent/.style={draw,-latex}]
      \node[draw,circle] {$W$}
      child {node[draw,circle] {$F$}}
      child {node[draw,circle] {$T$}};
    \end{tikzpicture}
  \end{subfigure}%
  \begin{subfigure}{0.8\textwidth}
    \centering
    \begin{tabular}[t]{cc}
      \toprule
      $w$ & $\Pr(W = w)$ \\
      \midrule
      1 & 0.5 \\
      0 & 0.5 \\
      \bottomrule
    \end{tabular}
    \begin{tabular}[t]{ccc}
      \toprule
      $w$ & $f$ & $\Pr(F = f \mid W = w)$ \\
      \midrule
      1 & 1 & 0.6 \\
      1 & 0 & 0.4 \\
      0 & 1 & 0.1 \\
      0 & 0 & 0.9 \\
      \bottomrule
    \end{tabular}
    \begin{tabular}[t]{ccc}
      \toprule
      $w$ & $t$ & $\Pr(T = t \mid W = w)$ \\
      \midrule
      1 & $l$ & 0.2 \\
      1 & $m$ & 0.4 \\
      1 & $h$ & 0.4 \\
      0 & $l$ & 0.6 \\
      0 & $m$ & 0.3 \\
      0 & $h$ & 0.1 \\
      \bottomrule
    \end{tabular}
  \end{subfigure}
  \caption{An example Bayesian network with its CPTs}
  \label{fig:example_bn}
\end{figure}

The Bayesian network in \cref{fig:example_bn} has
\begin{align*}
  V &= \{ W, F, T \}, \\
  \mathrm{pa}(W) &= \emptyset, \\
  \mathrm{pa}(F) &= \mathrm{pa}(T) = \{ W \}, \\
  \im W &= \im F = \{ 0, 1 \}, \\
  \im T &= \{ l, m, h \}, \\
  E(W) &= \{ \lambda_{W=1} \}, \\
  E(F) &= \{ \lambda_{F=1} \}, \\
  E(T) &= \{ \lambda_{T=l}, \lambda_{T=m}, \lambda_{T=h} \}, \\
  E^*(W) &= \{ \lambda_{W=1} \}, \\
  E^*(F) &= \{ \lambda_{F=1}, \lambda_{W=1} \}, \\
  E^*(T) &= \{ \lambda_{T=l}, \lambda_{T=m}, \lambda_{T=h}, \lambda_{W=1} \}, \\
  \mathrm{CPT_W} &= 0.5[\lambda_{W=1}]+0.5\overline{[\lambda_{W=1}]} = 0.5 \cdot 1, \\
  \mathrm{CPT_F} &= 0.6[\lambda_{F=1}] \cdot [\lambda_{W=1}] + 0.4[\lambda_{F=0}] \cdot [\lambda_{W=1}] + 0.1[\lambda_{F=1}] \cdot [\lambda_{W=0}] + 0.9[\lambda_{F=0}] \cdot [\lambda_{W=0}] \\
    &= 0.6[\lambda_{F=1}] \cdot [\lambda_{W=1}] + 0.4\overline{[\lambda_{F=1}]} \cdot [\lambda_{W=1}] + 0.1[\lambda_{F=1}] \cdot \overline{[\lambda_{W=1}]} + 0.9\overline{[\lambda_{F=1}]} \cdot \overline{[\lambda_{W=1}]}, \\
  \mathrm{CPT_T} &= ([\lambda_{T=l}] + [\lambda_{T=m}] + [\lambda_{T=h}]) \cdot (\overline{[\lambda_{T=l}]} + \overline{[\lambda_{T=m}]}) \cdot (\overline{[\lambda_{T=l}]} + \overline{[\lambda_{T=h}]}) \cdot (\overline{[\lambda_{T=m}]} + \overline{[\lambda_{T=h}]}) \cdot (\dots),
\end{align*}
and can be represented in a textual format as
\[
  \begin{array}{l r r l l}
    \lambda\sb{T=l} &\lambda\sb{T=m} &\lambda\sb{T=h} & &0 \\
                    &-\lambda\sb{T=l} &-\lambda\sb{T=m} & &0 \\
                    &-\lambda\sb{T=l} &-\lambda\sb{T=h} & &0 \\
                    &-\lambda\sb{T=m} &-\lambda\sb{T=h} & &0 \\
    w &\lambda\sb{W=1} & &0.5 &0.5 \\
    w &\lambda\sb{F=1} &\lambda\sb{W=1} &0.6 &0.4 \\
    w &\lambda\sb{F=1} &-\lambda\sb{W=1} &0.1 &0.9 \\
    w &\lambda\sb{T=l} &\lambda\sb{W=1} &0.2 &1 \\
    w &\lambda\sb{T=m} &\lambda\sb{W=1} &0.4 &1 \\
    w &\lambda\sb{T=h} &\lambda\sb{W=1} &0.4 &1 \\
    w &\lambda\sb{T=l} &\lambda\sb{W=0} &0.6 &1 \\
    w &\lambda\sb{T=m} &\lambda\sb{W=0} &0.3 &1 \\
    w &\lambda\sb{T=h} &\lambda\sb{W=0} &0.1 &1
  \end{array}
\]
with each $\lambda$ replaced with a unique positive integer. This format is
based on the format used by the Cachet solver \cite{DBLP:conf/sat/SangBBKP04}
to encode WMC problems, which extends the DIMACS format for CNF formulas with
weight clauses. Subsequently, we extend it in two ways:
\begin{itemize}
\item a single weight clause now supports an arbitrary number of literals,
\item and each weight clause has two probabilities instead of one (i.e., we no
  longer assume that $\Pr(v) + \Pr(\neg v) = 1$ for all variables $v \in U$).
\end{itemize}
The way we use this encoding, it is always the case that either both
probabilities sum to one, or the second probability (i.e., the probability for
the complement of the variable) is equal to one.

\subsection{Adjusting ADDMC to Work with the New Format}

ADDMC constructs the \emph{Gaifman graph} of the input CNF formula as an aid for
the algorithm's heuristics. This graph has as vertices the variables of the
formula, and there is an edge between two variables $u$ and $v$ if there is a
clause in the formula that contains both $u$ and $v$. We extend this definition
to functions on Boolean algebras, i.e., the factors of $\phi$. For any pair of
distinct variables $u, v \in U$, we draw an edge between them in the Gaifman
graph if there is a function $A\colon 2^X \to \mathbb{R}_{\ge 0}$ that is a
factor of $\phi$ such that $u \in X$ and $v \in X$. For instance, a factor such
as $\mathrm{CPT}_X$ will enable edges between all distinct pairs of variables in
$E^*(X)$.
